Complex Analysis of Askaryan Radiation: UHE-$\nu$ Identification and Reconstruction via the Hilbert Envelope of Observed Signals

### 2025-08-20 template.tex
updated unit 0 lecture notes, warm ups

\begin{abstract}
The detection of ultra-high energy neutrinos (UHE-$\nu$), with enegies above 10 PeV, has been a long-time goal of the astroparticle physics community.  A key strategy has been to deploy autonomous, radio-frequency (RF) detetectors in polar regions that rely on the Askaryan effect in ice for the neutrino signal.  The Askaryan effect occurs when the excess negative charge within a high-energy cascade radiates in a dense medium, and UHE-$\nu$ can induce such cascades that radiate in the radio-frequency (RF) bandwidth above thermal backgrounds.  To identify UHE-$\nu$ signals in future data from Askaryan-class detectors, analytic models of the Askaryan electromagnetic field have been created and compared to simulations and laboratory measurements.  These models have correctly described the Askaryan electromagnetic field, while leaving the effect of the RF detection channels on the field to simulation packages.  In this work, we present a fully analytic Askaryan model that accounts for the effect of an RF detection channel.  First, we derive formulas for the observed voltage trace, and for the Hilbert envelope of the trace.  Second, we match the analytic model to data computed for a detector with a single string of 8 RF detection channels in NuRadioMC, a key Monte Carlo toolset in UHE-$\nu$ detection.  Third, we find that a correlation threshold of 0.438 accepts 92\% of UHE-$\nu$ signals, and just 0.95 out of $1.57 \times 10^8$ RF thermal triggers.  At a thermal trigger rate of 1 Hz, this corresponds to 5 years of continuous operation.  Finally, we reconstruct the logarithm of the UHE-$\nu$ cascade energy with a precision of yyy.
\end{abstract}

@@

The constants $c_{\rm em}$ and $c_{\rm had}$ were found to be 0.80 and 0.93 meters, respectively, using Eqs. 10 and 12 from HH (\cite{PhysRevD.105.123019}), at FWHM for the high-energy cascades ($R=0.5$).  Using Eq. \ref{eq:lnLambda}, the $\sigma_t$ results from the optimized envelope fits to UHE-$\nu$ CSWs may be used to deduce the UHE-$\nu$ energy distribution.

>>>>

\begin{abstract}
The detection of ultra-high energy neutrinos (UHE-$\nu$), with enegies above 10 PeV, has been a long-time goal in astroparticle physics.  Autonomous, radio-frequency (RF) UHE-$\nu$ detetectors have been deployed in polar regions.  These detectors rely on the Askaryan effect in ice for the neutrino signal.  The Askaryan effect occurs when the excess negative charge within a high-energy cascade radiates in a dense medium.  UHE-$\nu$ can induce such cascades that radiate in the RF bandwidth above thermal backgrounds.  To identify UHE-$\nu$ signals in future data from Askaryan-class detectors, analytic models of the Askaryan electromagnetic field have been created and compared to simulations and laboratory measurements.  These models have correctly described the Askaryan electromagnetic field, while leaving the effect of the RF detection channels on the field to simulation packages.  In this work, we present a fully analytic Askaryan model that accounts for the effect of an RF detection channel.  First, we derive formulas for the observed voltage trace, and for the Hilbert envelope of the trace.  Second, we match the analytic model to data computed for a detector with a single string of eight RF detection channels in NuRadioMC, a key Monte Carlo toolset in UHE-$\nu$ detection.  For UHE-$\nu$ signals above RF thermal backgrounds, we find correlation coefficients in excess of $0.95$.  Third, 92\% of UHE-$\nu$ signals pass a correlation threshold of 0.438, while all but 0.95 out of $1.57 \times 10^8$ RF thermal triggers are rejected.  At a thermal trigger rate of 1 Hz, this corresponds to 5 years of continuous operation.  Finally, we reconstruct the logarithm of the UHE-$\nu$ cascade energy.
\end{abstract}

@@

Using Eqs. 10 and 12 from HH (\cite{PhysRevD.105.123019}), $c_{\rm em}$ and $c_{\rm had}$ were found to be 0.80 and 0.93 meters, respectively (FWHM, $R=0.5$).  Using Eq. \ref{eq:lnLambda}, the $\sigma_t$ results from the optimized envelope fits to UHE-$\nu$ CSWs may be used to deduce the UHE-$\nu$ energy distribution.  Two modifications facilitate interpretation.  First, converting to the base-10 logarithm is more convenient, and this introduces a factor of $\ln(10)$ in the denominator of Eq. \ref{eq:lnLambda}.  Second, $\ln\Lambda = \ln(E_{\rm C}/E_{\rm crit})$, where $E_{\rm C}$ is the energy in the UHE-$\nu$ induced cascade, and $E_{\rm crit}\approx 10^8$ eV is known as the critical energy \cite{PhysRevD.105.123019}.  Since $\ln\Lambda = \ln E_{\rm C} - \ln E_{\rm crit}$, separating this ratio simply adds a constant to the right hand side of Eq. \ref{eq:lnLambda}.  The modified form of Eq. \ref{eq:lnLambda} is

\begin{equation}
\log_{10} E_{\rm C} = \frac{\left( \frac{c\sigma_t}{c_{\rm em/had} \Delta\theta_{\rm rms}\sin\theta_{\rm C}} \right)^2}{\ln 10} + \log_{10} E_{\rm crit} \label{eq:lnLambda_2}
\end{equation}

Figure \ref{fig:fig6} contains the $\sigma_t$ distribution, converted into $\log_{10} E_{\rm C}$.  Note that when applying the change of base formula, factors of $\ln(10)$ cancel in the ratio of Eq. \ref{eq:a_err_2}.  The histogram is spread out over a range much larger than the expected range from NuRadioMC.  However, the average value of the histogram is $15.7 \pm 4.5$, in agreement with the prediction from NuRadioMC and Eq. \ref{eq:a_err_2}: $16.7 \pm 2.0$.  Thus, the energy reconstruction produces the correct order of magnitude on average, if not the precise value.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,trim=3.25cm 8.25cm 4.5cm 8.25cm,clip=true]{Aug19_plot1.pdf}
\caption{\label{fig:fig6} Normalized distribution of $\sigma_t$ values, converted to $\log_{10}E_{\rm C}$ via Eq. \ref{eq:lnLambda_2}.  The vertical black line represents the average $\log_{10}E_{\rm C}$ from the NuRadioMC energies, and the shaded region represnts the $1\sigma$ error range using Eq. \ref{eq:a_err_2}.}
\end{figure}
